Hierarchically modelling \emph{Kepler} dwarfs and subgiants to improve inference of stellar properties with asteroseismology

### 2020-11-05 paper/main.tex
Fixed typos and TFP comment

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[A. J. Lyttle et al.]{
Alexander J. Lyttle,$^{1,2}$\thanks{E-mail: ajl573@student.bham.ac.uk}
Guy R. Davies,$^{1,2}$
Tanda Li,$^{1,2}$
Lindsey M. Carboneau,$^{1,2}$
\newauthor
Oliver J. Hall,
Hin Leung,
Harry Westwood,
William J. Chaplin,$^{1,2}$
Daniel Huber,
\newauthor
Andrea Miglio,$^{1,2}$
Martin B. Nielsen,$^{1,2}$
and To B. Confirmed
\\
% List of institutions
$^{1}$School of Physics and Astronomy, University of Birmingham, Birmingham, B15 2TT, UK\\
$^{2}$Stellar Astrophysics Centre (SAC), Department of Physics and Astronomy, Aarhus University, Ny Munkegade 120, DK-8000 Aarhus C, Denmark\\
}

@@

% PARAGRAPH OUTLINE OF PAPER.
In Section \ref{sec:data}, we describe the data for the sample of 81 \emph{Kepler} dwarfs and subgiants studied in this work. We then present the methods in Section \ref{sec:meth} for which we produced a large grid of stellar evolutionary models to use as training data for an ANN (see Sections \ref{sec:grid} and \ref{sec:ann} respectively). We use the ANN as an emulator in a series of statistical models described in Section \ref{sec:hbm} to model the sample and present our results in Section \ref{sec:res}. Finally, in Section \ref{sec:dis} we discuss the results from each model and compare them to results for the sample sample of stars in the literature.

@@

We obtained results for each of the models described above by sampling their posterior distributions using a Markov chain Monte Carlo (MCMC) algorithm. In particular, we used the NUTS algorithm implemented in \textsc{PyMC4} -- a new version of \textsc{PyMC3} \citep{Salvatier.Wiecki.ea2016} based on \textsc{TensorFlow} \citep{Abadi.Barham.ea2016}. For each model, we produced \num{20000} samples split across \num{10} MCMC chains and computed summary statistics for the marginalised posteriors of each parameter in the model. We removed stars with problems during tuning using the Gelman-Rubin diagnostic \citep[$\hat{r}$;][]{Gelman.Rubin1992}. We used results from each model once $\hat{r} < 1.04$ for all parameters, indicating good model convergence.

Initially, we created a random synthetic population of stars using MESA to test ability of the method to recover stellar properties according to our choice of model physics and population priors. We tested the NP, PP and MP models. Since our sample was fictitious, it would not have been appropriate to include real solar data. We summarise the results for the synthetic stellar parameters and hyperparameters in Appendix \ref{sec:test-stars}. We found that the models were able to recover the true synthetic properties accurately, with increased precision when pooling parameters.

@@

During an initial run of the NP model, we dropped 16 of the 81 stars from the sample. Of the remove stars, we found the posteriors in $M$ for 6 skewed towards the prior upper mass limit of \SI{1.2}{\solarmass}. The remaining 10 removed stars suffered poor convergence during sampling ($\hat{r} >> 1.04$) which could be because of poor step-size tuning and sampling at the prior boundary.

@@

Out of the 65 stars with results from the NP model, 2 stars were dropped from the PP model with problems during sampling. A consequence of partially pooling parameters is that a population spread, $\sigma$ allows for individual parameters to vary outside of the prior range given to the population mean, $\mu$. In this case, individual stellar $\mlt$ was allowed to vary outside the range for which the ANN was valid if $\sigma_\alpha$ was large. The 2 removed stars happened to have high likelihoods outside of the $\mlt$ range which would be unphysical\footnote{We could truncate the population spread prior to avoid this problem, but found that removing the same 2 stars from the other models made negligible difference to the results, so we leave this to future work.}. In Tables \ref{tab:pp} and \ref{tab:pps} we present the results for the remaining 63 stars from the PP and PPS model respectively.

@@

To assess the accuracy of our model with respect to the literature, we compare our results to those of \citetalias{Serenelli.Johnson.ea2017} in Section \ref{sec:comp}. We found good agreement between the this work and their results, despite some differences in observables and stellar model physics which we discuss further.

@@

We found a greater difference in $\mlt$ between the models with and without the Sun when we max-pooled $\mlt$. The MP models yielded a global $\mlt$ in-line with $\mu_\alpha$ from the PP model. However, when we added the Sun, the model yields=ed $\mlt \approx 2.1$ which is in common with the solar results (see Appendix \ref{sec:sun-res}). This had a similar affect as assuming a solar calibrated value, because the model favoured fitting to data with the best observational precision. The change in $\mlt$ between the MP and MPS models resulted in a mean difference of $\sim 20$ per cent between the individual stellar ages. This is an example of how adopting a solar calibrated value can bias stellar ages. We argue that carefully including the Sun as a part of the population with an intrinsic spread is a better way to calibrate the stellar models.

@@

In Figure \ref{fig:unc-comp} we compare our statistical uncertainties with results for the equivalent stars from \citetalias{Serenelli.Johnson.ea2017} for mass, radius and age. We found that the NP model yielded comparable uncertainties to \citetalias{Serenelli.Johnson.ea2017} but note that these are likely underestimated due the influence of the prior boundaries for $Y_\mathrm{init}$ and $\mlt$. We expected larger uncertainties because we included additional free parameters ($Y_\mathrm{init}$ and $\mlt$) over the work of \citetalias{Serenelli.Johnson.ea2017}. However, when we treat these parameters hierarchically, we saw a reduction in uncertainties from all of the pooled models. This is because our prior assumptions about the population allows for the sharing of information between the stars. This uncertainty reduction scales with the number of stars in our sample, demonstrate by the results for the synthetic stars in Figure \ref{fig:shrinkage}. Thus, hierarchically modelling our population resulted in improved statistical uncertainties in stellar fundamental parameters.

>>>>

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[A. J. Lyttle et al.]{
Alexander J. Lyttle,$^{1,2}$\thanks{E-mail: ajl573@student.bham.ac.uk}
Guy R. Davies,$^{1,2}$
Tanda Li,$^{1,2}$
Lindsey M. Carboneau,$^{1,2}$
\newauthor
Oliver J. Hall,$^{1,2}$
Hin Leung,$^{1,2, 3}$
Harry Westwood,$^{1,2}$
William J. Chaplin,$^{1,2}$
Daniel Huber,
\newauthor
Andrea Miglio,$^{1,2}$
Martin B. Nielsen,$^{1,2}$
and To B. Confirmed
\\
% List of institutions
$^{1}$School of Physics and Astronomy, University of Birmingham, Birmingham, B15 2TT, UK\\
$^{2}$Stellar Astrophysics Centre (SAC), Department of Physics and Astronomy, Aarhus University, Ny Munkegade 120, DK-8000 Aarhus C, Denmark\\
$^{3}$School of Physics and Astronomy, University of St Andrews, North Haugh, St Andrews, KY16 9SS, UK\\
}

@@

% PARAGRAPH OUTLINE OF PAPER.
In Section \ref{sec:data}, we describe the data for the sample of 81 \emph{Kepler} dwarfs and subgiants studied in this work. We then present the methods in Section \ref{sec:meth} for which we produced a large grid of stellar evolutionary models to use as training data for an ANN (see Sections \ref{sec:grid} and \ref{sec:ann} respectively). We use the ANN as an emulator in a series of statistical models described in Section \ref{sec:hbm} to model the sample and present our results in Section \ref{sec:res}. Finally, in Section \ref{sec:dis} we discuss the results from each model and compare them to results for the sample of stars in the literature.

@@

We obtained results for each of the models described above by sampling their posterior distributions using a Markov chain Monte Carlo (MCMC) algorithm. In particular, we used the NUTS algorithm implemented in \textsc{TensorFlow Probability} \citep[\textsc{TFP};][]{Abadi.Barham.ea2016, Dillon.Langmore.ea2017}\footnote{We interacted with the \textsc{TFP} using the now depreciated \textsc{PyMC4} package, developed as a successor to \textsc{PyMC3} \citep{Salvatier.Wiecki.ea2016}}. For each model, we produced \num{20000} samples split across \num{10} MCMC chains and computed summary statistics for the marginalised posteriors of each parameter in the model. We removed stars with problems during tuning using the Gelman-Rubin diagnostic \citep[$\hat{r}$;][]{Gelman.Rubin1992}. We used results from each model once $\hat{r} < 1.04$ for all parameters, indicating good model convergence.

Initially, we created a random synthetic population of stars using MESA to test the ability of the method to recover stellar properties according to our choice of model physics and population priors. We tested the NP, PP and MP models. Since our sample was fictitious, it would not have been appropriate to include real solar data. We summarise the results for the synthetic stellar parameters and hyperparameters in Appendix \ref{sec:test-stars}. We found that the models were able to recover the true synthetic properties accurately, with increased precision when pooling parameters.

@@

During an initial run of the NP model, we dropped 16 of the 81 stars from the sample. Of the removed stars, we found the posteriors in $M$ for 6 skewed towards the prior upper mass limit of \SI{1.2}{\solarmass}. The remaining 10 removed stars suffered poor convergence during sampling ($\hat{r} >> 1.04$) which could be because of poor step-size tuning and sampling at the prior boundary.

@@

Out of the 65 stars with results from the NP model, 2 stars were dropped from the PP model with problems during sampling. A consequence of partially pooling parameters is that a population spread, $\sigma$ allows for individual parameters to vary outside of the prior range given to the population mean, $\mu$. In this case, individual stellar $\mlt$ was allowed to vary outside the range for which the ANN was valid if $\sigma_\alpha$ was large. The 2 removed stars happened to have high likelihoods outside of the valid $\mlt$ range. We found that removing the same 2 stars from the other models made negligible difference to the results, so we leave a solution to this problem to future work. In Tables \ref{tab:pp} and \ref{tab:pps} we present the results for the remaining 63 stars from the PP and PPS model respectively.

@@

To assess the accuracy of our model with respect to the literature, we compare our results to those of \citetalias{Serenelli.Johnson.ea2017} in Section \ref{sec:comp}. We found good agreement between this work and their results, despite some differences in observables and stellar model physics which we discuss further.

@@

We found a greater difference in $\mlt$ between the models with and without the Sun when we max-pooled $\mlt$. The MP models yielded a global $\mlt$ in-line with $\mu_\alpha$ from the PP model. However, when we added the Sun, the model yielded $\mlt \approx 2.1$ which is in common with the solar results (see Appendix \ref{sec:sun-res}). This had a similar affect as assuming a solar calibrated value, because the model favoured fitting to data with the best observational precision. The change in $\mlt$ between the MP and MPS models resulted in a mean difference of $\sim 20$ per cent between the individual stellar ages. This is an example of how adopting a solar calibrated value can bias stellar ages. We argue that carefully including the Sun as a part of the population with an intrinsic spread is a better way to calibrate the stellar models.

@@

In Figure \ref{fig:unc-comp} we compare our statistical uncertainties with results for the equivalent stars from \citetalias{Serenelli.Johnson.ea2017} for mass, radius and age. We found that the NP model yielded comparable uncertainties to \citetalias{Serenelli.Johnson.ea2017} but note that these are likely underestimated due the influence of the prior boundaries for $Y_\mathrm{init}$ and $\mlt$. We expected larger uncertainties because we included additional free parameters ($Y_\mathrm{init}$ and $\mlt$) over the work of \citetalias{Serenelli.Johnson.ea2017}. However, when we treat these parameters hierarchically, we saw a reduction in uncertainties from all of the pooled models. This is because our prior assumptions about the population allows for the sharing of information between the stars. This uncertainty reduction scales with the number of stars in our sample, demonstrated by the results for the synthetic stars in Figure \ref{fig:shrinkage}. Thus, hierarchically modelling our population resulted in improved statistical uncertainties in stellar fundamental parameters.
